The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-07-20 manuscript.Rmd
more changes

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood [@Worm2006], and modeling the impact of human activities has been limited to a few taxonomic groups [@Tittensor2010]. For marine organisms, because of the diversity of their complex life history strategies, their size range, and the difficulties associated with sampling their habitats, we have limited information about which species live where, or what portion of their diversity remains to be discovered and described [@Mora2011]. Additionally, coordinated marine ecosystem monitoring initiatives are lacking [@Duffy2013,@Borja2016], making it difficult to monitor changes in species distributions and abundances. While most documented global extinctions of marine species are vertebrates, many others species may have been lost unnoticed because they were not known [@Sala2006]. There is therefore a pressing need to improve our knowledge of marine systems before their functioning are irreversibly altered, and species go extinct.

@@

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), difficult to access, and therefore difficult to synthesize. The development of public databases for natural history records (e.g. iDigBio [@Page2015]), species occurrences (e.g., OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), now allow us to evaluate what and how much we know. Here, we aggregate publicly available data for marine invertebrates that occur within the limits of the exclusive economic zone (EEZ) of the continental United States (US), as well as the Gulf of Mexico and the Pacific Northwest, to highlight gaps in our knowledge that provide a road map for important new directions in global research. We show that while public databases are capturing a significant portion of our knowledge about species richness, most species are known from only a few specimens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but more comprehensive data on species distributions and abundances are needed. These goals can only be achieved with additional cost-effective sampling and a commitment to training more taxonomists and experts that can identify and interpret the material collected.

@@

To compare the faunistic lists from the Gulf of Mexico and the Pacific Northwest with records in public databases, we queried the iDigBio (https://portal.idigbio.org) and OBIS (http://www.iobis.org/) databases for all records they included in the same geographic boundaries as the lists. We filtered the results to only include the same higher taxonomic level taxa as in the faunistic lists, and kept only records for which the species could be matched to a species name in WoRMS. For the Gulf of Mexico, we retrieved  `r format(dplyr::n_distinct(idigbio_gom_records$worms_valid_name), big.mark=",")` species in iDigBio (from $n =$  `r format(nrow(idigbio_gom_records), big.mark=",")` records), and `r format(dplyr::n_distinct(obis_gom_records$worms_valid_name), big.mark=",")` species in OBIS ($n=$ `r format(nrow(obis_gom_records), big.mark=",")` records). For the Pacific Northwest, we retrieved `r format(dplyr::n_distinct(idigbio_pnw_records$worms_valid_name), big.mark=",")` species in iDigBio ($n=$ and `r format(nrow(idigbio_pnw_records), big.mark=",")`) and `r format(dplyr::n_distinct(obis_kozloff_records$worms_valid_name), big.mark=",")` species in OBIS ($n=$ `r format(nrow(obis_kozloff_records), big.mark=",")`). Overall, the number of species known from the Gulf of Mexico match the number estimated by experts (`r format(11150-1541-434, big.mark=",")` species [@Fautin2010]). For the Pacific coast, this number is lower than the estimate (`r format(8374-909-134, big.mark=",")` species [@Fautin2010]), but this estimate included species from California and from deeper areas that were not included in Kozloff's faunistic list.

@@

```{r idigbio-through-time, fig.cap='Cumulative number of samples (A) and species (B) of marine invertebrates recorded through time based on iDigBio records. Relation between number of samples collected, and number of species recorded for the first time (C), each point represents numbers for a given year. (\\#fig:idigbio-through-time)'}
knitr::include_graphics("figures/idigbio_through_time.pdf")
```

>>>>

With increasing impact of human activities on the biosphere, we need efficient and accurate approaches to document changes in species ranges and community composition. While terrestrial ecosystems have received considerable attention, the baseline of healthy and resilient marine ecosystems are not well understood [@Worm2006], and modeling the impact of human activities has been limited to a few taxonomic groups [@Tittensor2010]. For marine organisms, because of the diversity of their complex life history strategies, their size range, and the difficulties associated with sampling their habitats, we have limited information about which species live where, or what portion of their diversity remains to be discovered and described [@Mora2011]. Additionally, coordinated marine ecosystem monitoring initiatives are lacking [@Duffy2013,@Borja2016] (but see MarineGEO and PISCO), making it difficult to monitor changes in species distributions and abundances. While most documented global extinctions of marine species are vertebrates, many others species may have been lost unnoticed because they were not known [@Sala2006]. There is therefore a pressing need to improve our knowledge of marine systems before their functioning are irreversibly altered, and species go extinct.

@@

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), difficult to access, and therefore difficult to synthesize. The development of public databases for specimen records (e.g. iDigBio [@Page2015]), species occurrences (e.g., OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), now allow us to evaluate what and how much we know. Here, we aggregate publicly available data for marine invertebrates that occur within the limits of the exclusive economic zone (EEZ) of the continental United States (US), as well as the Gulf of Mexico (GOM) and the Pacific Northwest (PNW), to highlight gaps in our knowledge that provide a road map for important new directions in global research. We show that while public databases are capturing a significant portion of our knowledge about species richness, most species are known from only a few specimens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on communities require not only to get a more accurate list of the species that inhabits the oceans, but more comprehensive data on species distributions and abundances are needed. These goals can only be achieved with additional cost-effective sampling and a commitment to training more taxonomists and experts that can identify and interpret the material collected.

@@

To compare the faunistic lists from the Gulf of Mexico and the Pacific Northwest with records in public databases, we queried the iDigBio (https://portal.idigbio.org) and OBIS (http://www.iobis.org/) databases for all records they included in the same geographic boundaries as the lists. We filtered the results to only include the same higher taxonomic level taxa as in the faunistic lists, and kept only records for which the species could be matched to a species name in WoRMS. For the Gulf of Mexico, we retrieved  `r format(dplyr::n_distinct(idigbio_gom_records$worms_valid_name), big.mark=",")` species in iDigBio (from $n =$  `r format(nrow(idigbio_gom_records), big.mark=",")` records), and `r format(dplyr::n_distinct(obis_gom_records$worms_valid_name), big.mark=",")` species in OBIS ($n=$ `r format(nrow(obis_gom_records), big.mark=",")` records). For the Pacific Northwest, we retrieved `r format(dplyr::n_distinct(idigbio_pnw_records$worms_valid_name), big.mark=",")` species in iDigBio ($n=$ and `r format(nrow(idigbio_pnw_records), big.mark=",")`) and `r format(dplyr::n_distinct(obis_pnw_records$worms_valid_name), big.mark=",")` species in OBIS ($n=$ `r format(nrow(obis_pnw_records), big.mark=",")`). Overall, the number of species known from the Gulf of Mexico match the number estimated by experts (`r format(11150-1541-434, big.mark=",")` species [@Fautin2010]). For the Pacific coast, this number is lower than the estimate (`r format(8374-909-134, big.mark=",")` species [@Fautin2010]), but this estimate included species from California and from deeper areas that were not included in Kozloff's faunistic list.

@@

```{r idigbio-through-time, fig.cap='Cumulative number of samples (A) and species (B) of marine invertebrates recorded through time based on iDigBio records. Relation between number of samples collected, and number of species recorded for the first time (C), each point represents numbers for a given year, dashed lines show the 1:1 and 1:100 ratios for number of species recorded for the first time and number of samples. (\\#fig:idigbio-through-time)'}
knitr::include_graphics("figures/idigbio_through_time.pdf")
```
